---
title: "R Notebook"
output: html_notebook
---

\########################################

# OSA Use Case

::: {#dataCleaning .PRDL style="color: black"}
## **Simple ETL process on a single Excel file**

### Libraries

```{r}
library(readxl)
library(dplyr)
library(visdat)
library(naniar)
library(tidyr)
library(writexl)
```

\

```{r}
# Clear the working space
rm(list=ls())

Input_file <- "Info_BDApnea_QuironMalaga.xlsx"
Output_file <- "OSA_DB_UPM.xlsx"

#Data_Directory <- "DATA\\"
Data_Directory <- "C:\\Users\\maron\\OneDrive\\02-Documents\\00.ETUDES\\00.ECOLE_D_INGE\\00.CYCLE_ING_FORMATION_INIT\\00.3EME_ANNEE_INIT\\00.A_COURS\\01.MLLD\\02.ACTIVITIES\\OSA_CASE_STUDY_DONE\\DATA\\"
```

-   Using readxl package to read an Excel file

-   Install the readxl package is nor already installed

    ```{r}
    library(readxl)

    df_tmp <- read_excel(paste(Data_Directory, Input_file, sep = ""))

    typeof(df_tmp)
    is.data.frame(df_tmp)
    ```

    NOTE:

    :   -   **df_tmp** is NOT only a data frame!

        -   use as.data.frame to avoid later problems\

```{r}
class(df_tmp)
df_tmp = as.data.frame(df_tmp)
class(df_tmp)
head(df_tmp, 7)
```

PREPARING THE DATA FRAME

:   -   Select only those columns (variables) we are going

    -   to work with: **Patient**, **Gender**, **IAH**, **Peso**, **Talla**, **Edad**, **PerCervical**

    -   You can work with **dplyr** for using select(**df_tmp**, **Patient**, **Gender**,...)

    <https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html>

    -   dplyr's basic set of tools to apply on data frames

```{r}
library(dplyr)

df_tmp1 <- select(df_tmp, Patient, Gender, IAH, Peso, Talla, Edad, PerCervical)
```

HOW TO MANAGE Non Available values !!!

:   -   To change **-1** values for **NA** in all columns (i.e. variables) you can install package **naniar** load it

    -   **note:** you can also see

    <https://cran.r-project.org/web/packages/naniar/vignettes/getting-started-w-naniar.html>

    -   BUT first you can visualize the issue with chr, num and NA using visdat
    -   

```{r}
library(visdat)
vis_dat(df_tmp1)
```

BEFORE CONTINUING EXECUTING NEXT lines of code

:   -   BE SURE TO FIX the Excel removing comments in Peso

    -   **##########################QUESTIONS: \###################**

    -   HOW MANY PATIENTS?

    -   HOW MANY MALE /FEMALE

```{r}
library(naniar)

# Now change -1 values for NA in all columns (i.e. variables)
df_tmp2 <- replace_with_na_all(df_tmp1,condition = ~.x == -1)
df_tmp3 <- df_tmp2
df_tmp4 <- df_tmp2
```

### Removing the comments in the columns peso

```{r}
df_tmp3$Peso <- as.numeric(df_tmp3$Peso)
df_tmp3$Peso[is.na(df_tmp3$Peso)] <- NA

df_tmp4$Peso <- as.numeric(df_tmp3$Peso)
df_tmp4$Peso[is.na(df_tmp4$Peso)] <- 45
```

Note :

:   -   Finally remove (drop out) all rows containing a NA at least one column

    -   You can use **tidyr**

    -   <https://blog.rstudio.com/2014/07/22/introducing-tidyr/>

    -   **tidyr** is a package that makes it easy to "tidy" your data.

    -   Tidy data is data that's easy to work with:

    -   it's easy to munge (with **dplyr**), visualise (with **ggplot2** or **ggvis**) and model (with R's hundreds of modelling packages). The two most important properties of tidy data are:

    -   Each column is a variable.

    -   Each row is an observation.

```{r}
library(tidyr)

df_final <- df_tmp3 %>% drop_na()
```

QUESTIONS:

:   **\##### AFTER CLEANING: \#####**

    -   HOW MANY PATIENTS?

    -   HOW MANY MALE /FEMALE?

```{r warning=TRUE, paged.print=FALSE}
summ = summary(df_final)
print(summ)

nb_patient = length(df_final$Patient)
print(nb_patient)

nb_male = length(df_final$Gender[df_final$Gender == "hombre"])
print(nb_male)

nb_female = length(df_final$Gender[df_final$Gender == "mujer"])
print(nb_female)
```

SAVING CLEAN DATA

:   -   Write the clean data into Output_file

    -   you can install **writexl** package

```{r}
library(writexl)

# You can change the names of the columns
df_final <- df_final %>% rename(Weight = Peso,
                                Height = Talla,
                                Age = Edad,
                                Cervical = PerCervical)

vis_dat(df_final)

write_xlsx(df_final,
           paste(Data_Directory, Output_file, sep = ""))
```

```{r}
mean_IAH_mal = mean(df_final$IAH[df_final$Gender == "hombre"])
mean_IAH_female = mean(df_final$IAH[df_final$Gender == "mujer"])

print(mean_IAH_female)
print(mean_IAH_mal)

df_numerique <- df_final %>%
  select_if(is.numeric)

print(cor(df_numerique))
```

Data analysis:

```{r}
# Distribution of IAH by gender
Genders <- df_final$Gender
IAHs <- df_final$IAH
maleIAHs <- IAHs[Genders =="hombre"]
femaleIAHs <- IAHs[Genders =="mujer"]
muMale <- mean(maleIAHs)
muFemale <- mean(femaleIAHs)
b <- min(c(maleIAHs,femaleIAHs)) - 0.001
e <- max(c(maleIAHs,femaleIAHs))
ax <- pretty(b:e, n = 12)
hgA <- hist(maleIAHs, breaks = ax)
hgB <- hist(femaleIAHs, breaks = ax)
c1 <- rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue")
c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")
plot(hgA,col=c1,xlab="IAH")
plot(hgB,col=c2,add=TRUE)
```

\
:::

::: {#Analysis .PRDL style="color : black"}
# Notes :
:::

::: {#DataAnalysis .PRDL style="color : black"}
To analyse the data we can start by seperating the data into to subset, male and female subset

Then we make an analysis of the mean, correlation of the

plot the histogram for the two subset to see what kind of distribution do we have : Poisson distribution / The maen and the median ae not very different

We need the distribution to be Gaussian, then we have to make a transformation of the random variable : We can use Log(but in this case we need to add 1 because the IAH can be 0) or to take the inverse of the random variable

\
To check if two rv come from the same distribution or not : In our case the two subset

If the distribution of one of them is Gaussian then we can use the T-test, oif not we have to find another type of teste

The test can help to check if a feature is discriminative or not

some useful function :

-   filter

-   %\>%

-   nrow

-   

In some

print the summary of the two subset
:::
